Semantic Preserving Bijective Mappings for Expressions involving Special Functions between Computer Algebra Systems and Document Preparation Systems

### 2018-11-29 ReviewerResponse.txt

In the following we respond specifically to the reviewers comments.

Reviewer 1:

As suggested, we made several changes to improve the structure of the paper.  
In summary, we have :
(1) added subsection headings "Different Sets of Defined Functions", 
"Positions of Branch Cuts", "Insufficient Semantic Information", and 
"Ambiguous Expressions" in Section 3;
(2) gave an overview of the concept of the translator in Section 4;
(3) removed unnecessary Algorithm 1 in Section 5 as suggested, because 
it was just a simplified pre-version of Algorithm 2;
(4) added the subsection heading "Problems with the Look Ahead Approach" in 
Section 5 for a better outline; and 
(5) further explained the different states of Algorithm 3 (now Algorithm 2) 
by referring to specific lines in the algorithm. 
We also explained how a power after a macro is handled in more detail and why 
optional parameters need to be handled differently than other arguments.

### 2018-11-29 main.tex
Include def for appropriate/inappropriate translations

This article presents a new approach for automatic \gls*{p2c} and vice versa conversions. Translations from presentational to computational (computational to presentational) systems are called forward (backward) translations. A forward translation is denoted with an arrow with the target system language above the arrow. For example,
\begin{equation*}
t \overset{\langMaple}{\mapsto} c,
\end{equation*}
where $t$ is an expression in the \LaTeX{} language and $c$ is an element of the \Maple{} language $\langMaple$. As we will see later in this article, we need to compare mathematical concepts between systems. This is impossible from a mathematical point of view. Consider the irrational mathematical constant $e$, known as Euler's number. The theoretical construct for this symbol cannot be mathematically equivalent to the value \verb|exp(1)| in \Maple, caused by computational and implementational limitations. Instead of using the term \textit{equivalent}, we introduce \textit{appropriate} and \textit{inappropriate} translations. We call a translation such as
\begin{equation}\label{eq:cos-def}
\verb|\cos@{z}| \overset{\langMaple}{\mapsto} \verb|cos(z)|
\end{equation}
\textit{appropriate}, while a translation such as
\begin{equation}
\verb|\cos@{z}| \overset{\langMaple}{\mapsto} \verb|sin(z)|
\end{equation}
is \textit{inappropriate}. Note that it is not always as easy as in this example to decide if a translation is appropriate or not. Therefore, this article also presents several validation techniques to automatically verify if a translation is appropriate or inappropriate. In addition to this terminology, we introduce \textit{direct translations}. Later in the paper, we will explain that a translation from one specific mathematical object to its \textit{appropriate} counterpart in the other system is not always possible. We call a translation to the \textit{appropriate} counterpart \textit{direct}. For example, the translation~(\ref{eq:cos-def}) is \textit{direct}, while a translation to the definition of the cosine function
\begin{equation*}
\verb|\cos@{z}| \overset{\langMaple}{\mapsto} \verb|(exp(I*z)+exp(-I*z))/2|
\end{equation*}
is not a \textit{direct} translation.
Note that partial results of this paper have been published in~\parencite{CICM:Paper}.

@@

\vspace{-0.2cm}
The position of the branch cut of the arccotangent function differs after translation~\eqref{eq:acot-alternatives}, which may lead to incorrect calculations later on. The alternative translations~\eqref{eq:acot-alternatives-1} and~\eqref{eq:acot-alternatives-2} use other functions instead of the arccotangent function. The arctangent function~\eqref{eq:acot-alternatives-1} and the natural logarithm~(\ref{eq:acot-alternatives-2}) have the same positioned branch cuts as in the \gls*{dlmf} and in \Maple. Hence, translation~(\ref{eq:acot-alternatives-1}) solves the issue, as long as the user does not evaulate the function at $z = 0$, while translation~(\ref{eq:acot-alternatives-2}) solves the issue except at $z = -\iunit$.

>>>>

This article presents a new approach for automatic \gls*{p2c} and vice versa conversions. Translations from presentational to computational (computational to presentational) systems are called forward (backward) translations. A forward translation is denoted with an arrow with the target system language above the arrow. For example,
\begin{equation*}
t \overset{\langMaple}{\mapsto} c,
\end{equation*}
where $t$ is an expression in the \LaTeX{} language and $c$ is an element of the \Maple{} language $\langMaple$. As we will see later in this article, we need to compare mathematical concepts between systems. This is impossible from a mathematical point of view. Consider the irrational mathematical constant $e$, known as Euler's number. The theoretical construct for this symbol cannot be mathematically equivalent to the value \verb|exp(1)| in \Maple, caused by computational and implementational limitations.

Instead of using the term \textit{equivalent}, 
we introduce the terms \textit{appropriate} and \textit{inappropriate} translations. 
%{\bf [HSC: I don't understand how {\it equivalent} translations has anything at 
%all to do with {\it appropriate} or {\it inapppropriate} translations. Please clarify.]}.
We consider a translation as {\it appropriate}, when a numerical evaluation returns the 
same values in both concepts up to a numerical precision $|\epsilon|<<1$, for all possible 
points in the domain of the functions.
A translation is considered as {\it inappropriate}, when it is not {\it appropriate}.

For example, a translation such as
\begin{equation}\label{eq:cos-def}
\verb|\cos@{z}| \overset{\langMaple}{\mapsto} \verb|cos(z)|
\end{equation}
is \textit{appropriate}, while a translation such as
\begin{equation}
\verb|\cos@{z}| \overset{\langMaple}{\mapsto} \verb|sin(z)|
\end{equation}
is \textit{inappropriate}. Note that it is not always as easy as in this example to decide if a translation is {\it appropriate} or not. Therefore, this article also presents several validation techniques to automatically verify if a translation is {\it appropriate} or {\it inappropriate}. In addition to this terminology, we introduce \textit{direct translations}. Most mathematical objects in one system have a direct counterpart in other systems. However, later in the paper, we will explain that a translation from one specific mathematical object to its counterpart in the other system is not always \textit{appropriate}. Also, not every mathematical object has a counterpart in other systems. We call a translation to its counterpart \textit{direct}. For example, the translation~(\ref{eq:cos-def}) is \textit{direct}, while a translation to the definition of the cosine function
\begin{equation*}
\verb|\cos@{z}| \overset{\langMaple}{\mapsto} \verb|(exp(I*z)+exp(-I*z))/2|
\end{equation*}
is not a \textit{direct} translation even though it is {\it appropriate}.
Note that partial results of this paper have been published in~\parencite{CICM:Paper}.

@@

\vspace{-0.2cm}
The position of the branch cut of the arccotangent function differs after the {\it direct} translation~\eqref{eq:acot-alternatives}, which may lead to incorrect calculations later on. The alternative translations~\eqref{eq:acot-alternatives-1} and~\eqref{eq:acot-alternatives-2} use other functions instead of the arccotangent function. The arctangent function~\eqref{eq:acot-alternatives-1} and the natural logarithm~(\ref{eq:acot-alternatives-2}) have the same positioned branch cuts as in the \gls*{dlmf} and in \Maple. Hence, translation~(\ref{eq:acot-alternatives-1}) solves the issue, as long as the user does not evaulate the function at $z = 0$, while translation~(\ref{eq:acot-alternatives-2}) solves the issue except at $z = -\iunit$. Note that none of the translations~(\ref{eq:acot-alternatives}-\ref{eq:acot-alternatives-2}) are {\it appropriate}.
